---
title: "R Notebook"
output: html_notebook
---

```{r message=FALSE}
library(tidyverse)
library(magrittr)
library(forcats)
```


```{r message=FALSE}
audits <- 
  list.files("../../scratch/CRISPR_PILOT_B1/audit/", pattern = "*_audit_detailed.csv", 
             full.names = TRUE) %>% 
  map_df(read_csv)
```


```{r}
audits %<>% 
  select(Metadata_gene_name,
         Metadata_pert_name,
         Metadata_broad_sample,
         Metadata_cell_line,
         correlation)

```


```{r}

audits %<>%
  group_by(Metadata_gene_name, Metadata_pert_name) %>%
  summarise_at("correlation", median) %>%
  ungroup() %>%
  group_by(Metadata_gene_name) %>%
  summarise_at("correlation", median) %>%
  ungroup() %>%
  rename(median_correlation = correlation) %>%
  inner_join(audits, by = c("Metadata_gene_name"))


audits %<>% 
  select(Metadata_gene_name, Metadata_pert_name, correlation) %>% 
  group_by(Metadata_gene_name, Metadata_pert_name) %>% 
  summarise_at("correlation", median) %>%
  ungroup() %>%  
  group_by(Metadata_gene_name) %>% 
  arrange(correlation) %>% 
  ungroup() %>%
  select(-correlation) %>%
  group_by(Metadata_gene_name) %>% 
  mutate(guide_index = row_number()) %>% 
  ungroup() %>% 
  inner_join(audits, by = c("Metadata_gene_name", "Metadata_pert_name"))
```


Plot replicate correlation for each guide, grouped by gene, faceted by cell line.
Order genes by the average replicate correlation across all guides. Each dot
corresponds to a well. The colors correspond to an index assigned to a 
guide for a gene (e.g. if a gene has two guides in this expt, they will have 
indices of 1 and 2, where 1 has a higher average replicate correlation)

```{r fig.width=7}

p <- 
ggplot(audits %>% filter(Metadata_gene_name != "Chr2"),
       aes(fct_reorder(Metadata_gene_name, median_correlation),
           correlation)) + 
  geom_point(aes(color = as.factor(guide_index))) + 
  facet_wrap(~Metadata_cell_line, ncol = 1) + 
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

p

ggsave("all_guides.pdf", width = 12, height = 6)

```

Filter down to only those guides that have a Broad ID
```{r fig.width=7}
p <- 
ggplot(audits %>% 
         filter(Metadata_gene_name != "Chr2") %>%
         filter(!is.na(Metadata_broad_sample)),
       aes(fct_reorder(Metadata_gene_name, median_correlation),
           correlation)) + 
  geom_point(aes(color = as.factor(guide_index))) + 
  facet_wrap(~Metadata_cell_line, ncol = 1) + 
  theme(axis.text.x=element_text(angle = 45, hjust = 1))

p

ggsave("only_guides_with_barcodes.pdf", width = 12, height = 6)

```

```{r}
write_csv(audits, "audits.csv")
```

