---
title: "CellHealth_vs_TA_ORF"
author: "TB"
date: "4/19/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R 

```{r load libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(reshape2)
library(tidyr)
library(tibble)
library(readr)
library(stringr)
library(ape)
library(ggplot2)
library(corrplot)
```

# Data
Load Cell Health data and TA ORF data. 

For TA ORF we load normalized data with selected features only 

For CellHealth we load normalized data 

```{r load data, message=FALSE, warning=FALSE}

read_profiles <- function(data_dir,csv_identifier ) {
  dir_list <- list.dirs(data_dir, recursive = FALSE)
  df_profiles <- c()

  for (dir_i in dir_list) {
    file_lst <- list.files(dir_i)
    file_name <- file_lst[str_detect(file_lst, csv_identifier)]
    df <- suppressWarnings(read_csv(sprintf("%s/%s",dir_i, file_name)))
    df_profiles <- c(df_profiles, list(df))
  }

  df_profiles <- do.call(rbind, df_profiles)
  colnames_profiles <- colnames(df_profiles)
  meta_cols <- colnames_profiles[str_detect(colnames_profiles, "Metadata_")]
  feat_cols <- setdiff(colnames_profiles, meta_cols)
  Pf.data <- list(data = df_profiles, feat_cols = feat_cols, meta_cols = meta_cols, backend = data_dir)
  return(Pf.data)
}

file_name_common_genes_collapsed <- '../../scratch/CRISPR_PILOT_B1/crispr_vs_orf/CRISPR_TAORF_common_genes_collapsed_normalized_variable_selected.txt'
file_name_common_genes <- '../../scratch/CRISPR_PILOT_B1/crispr_vs_orf/CRISPR_TAORF_common_genes_normalized_variable_selected.txt'


dir_cell_health <- "../../backend/CRISPR_PILOT_B1"
csv_identifier_cell_health <- "_normalized_variable_selected.csv"
Pf.cell_health <- read_profiles(dir_cell_health, csv_identifier_cell_health)
df_cell_health <- Pf.cell_health$data

dir_ta_orf <- "../../backend/SIGMA2_Pilot_2013_10_11"
csv_identifier_ta_orf <- "_normalized.csv"
Pf.ta_orf <- read_profiles(dir_ta_orf, csv_identifier_ta_orf)
df_ta_orf <- Pf.ta_orf$data

Pf.cell_health$data <- Pf.cell_health$data %>% 
  dplyr::mutate(Metadata_Treatment = paste(Metadata_pert_name, Metadata_cell_line, sep = "_")) %>%
  as.data.frame()

```
# select genes with a strong signal
Any genes is selected that shows a strong signal in at least one of the CRISPR cell lines (A549, ES2, HCC44) 

```{r select genes with signal}

median_pairwise_correlation <- function(df, variables, group_by) {
  df %>% 
    dplyr::group_by_(.dots = group_by) %>% 
    do(tibble::data_frame(correlation = median(as.dist(cor(t(as.matrix(.[variables])))))))
}


estimate_null_threshold <- function(df, feat.cl, group_by_variable, n_fold) {
  us <- c()
  u <-  c()
  for (i in 1:n_fold) {
    u <- df %>% 
    tidyr::unite_("group_by_variable", group_by_variable) %>%
    mutate(group_by_variable = sample(group_by_variable)) %>%
    median_pairwise_correlation(feat.cl, "group_by_variable") 
    us <- rbind(us,u)
  }
  null_threshold <- us %>%
    magrittr::extract2("correlation") %>%
    quantile(0.95, na.rm = TRUE) 
  
  return(null_threshold)
}


group_by_variable <- "Metadata_pert_name"

correlations <- Pf.cell_health$data %>% 
  median_pairwise_correlation(Pf.cell_health$feat_cols, group_by_variable) 

null_thresholds <- Pf.cell_health$data %>% group_by(Metadata_cell_line) %>% 
  do(data.frame(null_threshold = estimate_null_threshold(., Pf.cell_health$feat_cols, group_by_variable,10) ))

strong_genes <- Pf.cell_health$data %>%
  right_join(null_thresholds, by = 'Metadata_cell_line') %>%
  right_join(correlations, by = 'Metadata_pert_name') %>% 
  select(correlation, null_threshold, Metadata_cell_line, Metadata_pert_name, Metadata_gene_name ) %>% 
  filter(correlation > null_threshold) %>% 
  select(Metadata_gene_name) %>% 
  unique()

``` 


```{r, select common genes with a strong signal and features }
common_genes <- df_ta_orf %>% 
  select(Metadata_gene_name) %>%
  unique %>%
  intersect(.,df_cell_health %>% 
    select(Metadata_gene_name) %>%
    unique) %>%
  dplyr::filter(Metadata_gene_name != 'EMPTY' )

common_features <- intersect(Pf.ta_orf$feat_cols, Pf.cell_health$feat_cols) 

nan.col <- apply((df_ta_orf),2,function(x) return(all(!is.na(x))))
common_features <- setdiff(common_features, names(which(!nan.col)))

common_genes <- intersect(strong_genes, common_genes)


common_genes %>% 
  knitr::kable()


```


```{r BRAF }

combine_matrices <- function(df1, df2, gene_name, common_features) {
  df1 %>% 
    filter(Metadata_gene_name %in% gene_name ) %>%
    select( one_of(c(common_features, "Metadata_cell_line","Metadata_pert_name","Metadata_gene_name"))) %>% 
    arrange(Metadata_cell_line, Metadata_pert_name) %>% 
    mutate(Metadata_experiment = 'CRISPR') %>%
    rbind( df2   %>% 
      filter(Metadata_gene_name %in% gene_name ) %>%
      select( one_of(c(common_features, "Metadata_cell_line","Metadata_pert_name","Metadata_gene_name"))) %>% 
      arrange(Metadata_cell_line, Metadata_pert_name) %>% 
      mutate(Metadata_experiment = 'TA ORF') ) %>%
    mutate(perturbation = paste(Metadata_experiment, Metadata_cell_line, Metadata_pert_name, sep = "_"))
      
}

get_correlation_matrix <- function(df1, df2, gene_name, common_features){
  df <- combine_matrices(df1, df2, gene_name, common_features)
     
  cor_matrix <- df %>%
    select(one_of(common_features)) %>%
    t %>%
    cor
  
  labels <- df %>% select(perturbation)
  rownames(cor_matrix) <- labels[[1]]  
  
  return(cor_matrix)
}

for (i_gene_name in common_genes[[1]]) {
  cm <- get_correlation_matrix(df_cell_health, df_ta_orf, i_gene_name, common_features)
  corrplot(cm,title = i_gene_name, diag = TRUE, mar = c(0,0,1,0)) #, method = "pie")
}

cm <- get_correlation_matrix(df_cell_health, df_ta_orf, common_genes$Metadata_gene_name, common_features)
corrplot(cm,title = "common genes", diag = TRUE, mar = c(0,0,1,0))
hc <- hclust(dist(cm))
plot(as.phylo(hc), type = "fan",  label.offset = 0.3)

df <- combine_matrices(df_cell_health, df_ta_orf, common_genes$Metadata_gene_name, common_features)
df <- df %>% select(Metadata_cell_line, Metadata_pert_name, Metadata_gene_name, Metadata_experiment, everything() )


write.table(df, file = file_name_common_genes, quote = FALSE, sep = '\t', row.names = FALSE)

```
# collapsing the data
collapse all data for one gene. 
Question: which method to use? 
*median
*quantile? 

```{r collapsing the data}


collapse_profiles <- function(df, exp_label ) {
  df %>%  
  dplyr::filter(Metadata_gene_name %in% common_genes[[1]]) %>%
  mutate(Metadata_experiment = exp_label) %>%
  select(one_of(c(common_features, 'Metadata_experiment', 'Metadata_cell_line', 'Metadata_gene_name', 'Metadata_pert_name'))) %>%
  group_by(Metadata_experiment, Metadata_cell_line, Metadata_gene_name) %>% 
  summarise_each(funs(mean)) %>%
  return()
}

df_collapsed <- rbind( 
  collapse_profiles(df_cell_health,'CRISPR'),
  collapse_profiles(df_ta_orf,'TA ORF')
)

cor_matrix <- df_collapsed %>%
  ungroup %>% 
  select(one_of(common_features)) %>%
  t %>%
  cor
  
labels <- df_collapsed %>% 
  ungroup %>%
  select(Metadata_cell_line, Metadata_pert_name, Metadata_gene_name, Metadata_experiment, everything() ) %>%
  mutate(perturbation = paste(Metadata_experiment, Metadata_cell_line, Metadata_gene_name, sep = "_")) %>%
  select(perturbation)

rownames(cor_matrix) <- labels[[1]]  

cell_line_index <- df_collapsed %>% group_by(Metadata_cell_line) %>% group_indices() 
  
corrplot(cor_matrix,title = "22 collapsed genes", diag = TRUE, mar = c(0,0,1,0))

hc_abs <- hclust(dist(abs(cor_matrix)))
mypal = c( "#4ECDC4", "#556270", "#1B676B",  "#FF6B6B")
plot(as.phylo(hc_abs), type = "fan",  label.offset = 0.1,tip.color = mypal[cell_line_index] )
  
hc <- hclust(dist(cor_matrix))
mypal = c( "#4ECDC4", "#556270", "#1B676B",  "#FF6B6B")
plot(as.phylo(hc), type = "fan",  label.offset = 0.1,tip.color = mypal[cell_line_index] )

df_collapsed <- df_collapsed %>% select(Metadata_cell_line, Metadata_pert_name, Metadata_gene_name, Metadata_experiment, everything() )
write.table(df_collapsed, file = file_name_common_genes_collapsed, quote = FALSE, sep = '\t', row.names = FALSE)

```
# dot plot
```{r }

get_single_gene_cor <- function(df, g_name) {
  df_single_gene <- df %>% 
  ungroup %>% 
  filter(Metadata_gene_name == g_name) %>%
  mutate(Data = paste(Metadata_cell_line, Metadata_experiment )) 
  
  cor_labels <- df_single_gene %>%
    select(Metadata_cell_line)
  
  cor_matrix <- df_single_gene %>%
    select(one_of(common_features)) %>%
    t %>%
    cor
  
  rownames(cor_matrix) <- cor_labels[[1]] 
  
  return(cor_matrix)
}

n_genes <- common_genes %>% select(Metadata_gene_name) %>% summarise(n = n()) %>% print

cor_label <-  rep(c('A549 ES2', 'A549 HCC44', 'A549 U2OS', 'ES2 HCC44', 'ES2 U2OS', 'HCC44 U2OS'), n_genes)

genes <- rep(common_genes$Metadata_gene_name,each = 6)



cor_list <- c()
perturbation = rep(c(1,1,2,1,2,2),n_genes)

for (i_gene in as.vector(common_genes$Metadata_gene_name)) {
  i_gene %>% print
  cm <- get_single_gene_cor(df_collapsed, i_gene)  
  corrplot(cm, diag = TRUE, title = i_gene, mar = c(0,0,1,0))
  cor_list <- c(cor_list, c(cm[2:4,1],cm[3:4,2],cm[4,3]))
}

df_cor <- data.frame(correlation = as.vector(cor_list), cor_label, genes, perturbation)

corrplot(cm, diag = TRUE, title = common_genes$Metadata_gene_name[22], mar = c(0,0,1,0))

#ggplot(data = df_cor) + 
#  geom_point(mapping = aes(x = position, y = correlation), color = perturbation, shape = perturbation) 

```
```{r}

group.t.test <- function(df_cor.in) {
g1 <- df_cor.in %>%
  filter(perturbation == 1) %>%
  select(correlation) %>% 
  as.matrix() %>% 
  as.vector()

g2 <- df_cor.in %>%
  filter(perturbation == 2) %>%
  
  select(correlation) %>% 
  as.matrix() %>% 
  as.vector()
  
  return(t.test(g1, g2, alternative = "greater")$p.value)
}

p.vals <- df_cor %>%
  group_by(genes) %>%
  do(data.frame(p.value = group.t.test(.)))

p.vals$p.value <- p.adjust(p.vals$p.value, method = "BH")

p.vals  %>% arrange(p.value) %>% print 

sorted_index <- p.vals %>% as.data.frame()  %>% arrange(p.value)  %>% mutate(index = row_number()) %>% print 

df_cor_ext <- right_join(df_cor, sorted_index, by = "genes")


x <- df_cor_ext[,c("genes", "index")] %>% arrange(index)
df_cor_ext$genes <- factor(x$genes, levels = x$genes)

ggplot(data = df_cor_ext) + 
  geom_point(mapping = aes(x = genes, y = correlation), color = perturbation, shape = perturbation) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))

p.vals %>% arrange(p.value) %>% knitr::kable()
```

```{r}
Pf.cell_health$data %>% 
  filter(Metadata_cell_line == 'A549') %>% 
  select(Nuclei_AreaShape_MedianRadius, Metadata_pert_name) %>% 
  arrange(-Nuclei_AreaShape_MedianRadius) %>% 
  print
```
```{r, guides per gene}
df <- Pf.cell_health$data 
  
guides_per_gene <- function(df) {
  df %>% 
  group_by(Metadata_gene_name) %>% 
  summarize(n = length(unique(Metadata_pert_name))) %>%
  return
}

knitr::kable(guides_per_gene(df))
  
```